\section{Related Work}

\paragraph{Dense retrieval as bi-encoding.}
Dense passage retrieval is usually cast as a dual-encoder problem in which queries and passages are embedded independently and scored with an inner product or cosine similarity. DPR established that this simple formulation can replace sparse first-stage retrieval in open-domain QA, and Sentence-BERT showed that siamese and triplet-style text encoders make single-vector semantic search practical at scale \cite{karpukhin-etal-2020-dense,reimers-gurevych-2019-sentence}. Our study uses that standard bi-encoder setup and asks how much quality remains when only the query side is compressed.

\paragraph{Contrastive training and representation heads.}
In-batch contrastive learning is the default optimization recipe for dense retrievers, and SimCLR is the canonical reference for normalized contrastive objectives and projection heads \cite{chen-etal-2020-simclr}. Dense retrieval systems such as RocketQA show that performance is sensitive to negatives, batching, and training/inference mismatch, while recent distillation work such as PairDistill continues to refine teacher-student supervision for retrieval \cite{qu-etal-2021-rocketqa,huang-chen-2024-pairdistill}. These lines of work motivate our pooling, projection-head, and lightweight distillation ablations, but they are broader than the narrow asymmetric setting studied here.

\paragraph{Embedding teachers and zero-shot evaluation.}
Strong embedding models now serve as practical frozen references for retrieval studies. The official \texttt{BAAI/bge-base-en-v1.5} model card documents the retrieval instruction and CLS-based embedding usage that we follow for the teacher reference, and it reports strong zero-shot retrieval performance on standard benchmarks \cite{bge-base-en-v15-modelcard}. For transfer evaluation, BEIR remains the standard heterogeneous zero-shot benchmark and is well aligned with our emphasis on generalization rather than leaderboard fitting \cite{thakur-etal-2021-beir}.

\paragraph{ANN indexing.}
At query time, dense retrieval systems rely on exact or approximate nearest-neighbor search over a fixed vector index. Faiss is the standard library for this layer, and its flat, IVF, and HNSW families provide the canonical exact and approximate baselines for large-scale vector search \cite{johnson-etal-2021-faiss,malkov-yashunin-2018-hnsw}. In this paper, ANN is treated as a systems comparison layered on top of the retrieval model, not as a new retrieval objective.
