\section{Experimental Setup}

The benchmark uses an MS MARCO-derived slice prepared with seed $13$. The resulting training and retrieval files contain $100{,}000$ training triples, $351{,}695$ corpus passages, $55{,}578$ development queries, and $59{,}273$ qrels. Training and evaluation text are truncated to 256 tokens, and all query texts use the retrieval instruction prefix \texttt{Represent this sentence for searching relevant passages: } while passages use no prefix.

We report a frozen zero-shot teacher reference based on \texttt{BAAI/bge-base-en-v1.5} with CLS pooling. The main student family is the asymmetric 4-layer query encoder with mean pooling on the query side and CLS pooling on the passage side, which is the best-performing shallow student in the present study. Additional ablations include the raw 4-layer and 2-layer CLS students, a query-side projection-head variant, and a lightweight KL distillation variant. The teacher fine-tuning ablation is retained only as a comparison point; it does not replace the frozen zero-shot reference.

Training uses AdamW with batch size $16$, weight decay $0.01$, warmup ratio $0.1$, and fixed learning rates for each run family. The teacher and student runs use the same tokenizer and the same MS MARCO-derived triples, with validation performed on a deterministic 5,000-example slice of the training triples. Checkpoint selection is based on retrieval probe \texttt{MRR}@10 when a probe is enabled, which keeps the selection criterion aligned with retrieval quality rather than only validation loss.

Retrieval evaluation is performed on the full prepared corpus and development query set with FAISS exact inner-product search for the main tables. We report $\mathrm{MRR}@10$ and $\mathrm{Recall}@100$ together with corpus encoding time, index build time, query encoding latency, search latency, and end-to-end online latency. Latency is measured with batch size $1$ after a short warmup, using p50 and p95 over a fixed 100-query sample. Query-side memory is reported from the measured latency runs. The main quantitative summaries consist of a retrieval table, a latency table, an ANN table, and one latency--quality tradeoff plot.

For robustness, we also ran a BEIR transfer check on SciFact using the strongest shallow student. That transfer result is reported as an auxiliary sanity check rather than a primary target. The ANN extension compares exact FAISS against HNSW and IVF on the strongest shallow student, using the same corpus/query slice and the same retrieval metrics so that speed and quality are directly comparable.
